Notes of Basic Topolgy

### 2016-12-11 Basic Topology Notes/sections/ch6-Triangulation.tex
[major] Introduces Edge group in topology.

>>>>

\subsection{6.4 The edge group of a complex}
\label{sec:edge-group}

The definition of edge group is so similar to that of fundamental
group (actually, edge group is made to simplify computation of
fundamental group). Therefore, I will not waste too much words here.
One with reasonable intuition could just skim through this section.

The following excerpt from book \cite{book} best captures the idea:

\begin{myquote} \enquote{
We calculated the fundamental groups of one or two spaces in Chapter
5, but our calculations, though efficient for the examples given
there, were rather \textit{ad hoc}. If we agree to work with triangulable
spaces, we can be much more systematic. We shall show how to read off
generators and relationst for the fundamental group from a
triangulation of the space.

Let $X$ be a path-connected triangulable space, take a specific
triangulation $h:|K|\to X$, and replace $X$ by $|K|$ (we are at
liberty to do this since the funda- mental group is a topological
invariant). Now the advantage of a polyhedron $|K|$ is that the
elements of its fundamental group can be represented by loops which
are made up of edges of $K$. Using such 'edge loops' we shall
construct a group, called the edge group of the complex $|K|$, which
can be computed and wh ich is isomorphie to the fundamental group of
$|K|$.
} \end{myquote}

The edge pathes $v_0,v_1,\cdots,v_k$, edge loops
$v,v_1,\cdots,v_{k-1},v$ are defined similar to pathes and loops.
The equivalence relation between edge loops are defined as represented
by the following drawing:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/ch6-notes-2/edge-group1.png}
\end{figure}

The equivalence class of edge pathes and edge loops are denoted
$\{v_0,v_1,\cdots,v_k\}$, and $\{v,v_1,\cdots,v_{k-1},v\}$,
respectively. The multiplication of edge loops is defined simply as
\begin{equation}
    \{v,v_1,\cdots,v\}\{v,w_1,\cdots,v\} =
    \{v,v_1,\cdots,v,w_1,\cdots,v\}
\end{equation}

It is trivially verified that such edge loops form a group, called the
\nomen{edge group based at $v$, $E(K,v)$}. It is also trivially
verified that such group does not depend on the choice of base point
$v$, since $|K|$ is path-connected.

\begin{thm}
    $E(K,v)$ is isomorphie to $\pi_1(|K|,v)$
\end{thm}
\begin{proof}
    The proof can be found on page 133 of \cite{book}. The essential
    idea is that an edge loop is just a loop. And a loop is just a
    continuous map $f:I\to |K|$, we can easily use the simplicial
    approximation $s:I^m\to |K|$ of $f$ (with appropriately $m$), to
    find a corresponding edge loop.

    The uniqueness of the loop corresponding to an edge loop is found
    again by treating $L:=I^2$ as a simplicial complex and apprximate
    again the homotopy $F:L\to |K|$. The resulted $S$ will again gives
    us a equivalence relation to the constant edge loop. (Note, we
    have to use the fact the simplicial maps preserves the equivalence
    relation between edge pathes). The following pictorially
    produces the proof:
    \begin{figure}[H]
        \centering
        \includegraphics[width=1\linewidth]{pics/ch6-notes-2/L.png}
    \end{figure}
\end{proof}

### 2016-12-11 Basic Topology Notes/sections/ch6-Triangulation1.tex
[major] Introduces Edge group in topology.

\subsection{Section 6.2 Barycentric Division}
\label{sec:Barycentric-Division}
Now we begin to make the triangulation more and more detailed. We
start from a simplicial complex $K$ and constructing a new simplicial
complex $K_1$, such that
\begin{enumerate}
    \item $|K|=|K_1|$, homeomorphically.
    \item diameter $|K_1|<$ diameter $|K|$.
\end{enumerate}
The general construction method is called Barycentric division,
visualized as:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/ch6-scanned-notes-1/8.pdf}
\end{figure}
We first define a concept
\begin{defi}[$\hat{A}$ Bary center]
\nomenclature{$\hat{A}$ Bary center}{\nomrefpage.}
Let $A$ be a simplex, then the bery center of $A$, denoted $\hat{A}$,
is the point:
\begin{equation}
    \hat{A}:= \frac{1}{k+1}(v_0+\cdots+v_k)
\end{equation}
\end{defi}
Then, the general process is:
The new space $K_1$ is such that
\begin{enumerate}
    \item The vertices of $K_1$ are bary centers of \textbf{ALL}
        simplexes of $K$. Note the the bary center of a $0$-simplex
        is just itself. So in general, $K_1$ has more vertices of
        $K$.
    \item A collection $\hat{A}_0,\cdots,\hat{A}_k$ of such bary
        centers will form a vertex of a $k$-simplex in $K_1$ if and
        only if there is a permutation $\sigma$ of $\{0,1,\cdots,k\}$
        such that
        \begin{equation}
            A_{\sigma{0}}< A_{\sigma{1}} < \cdots < A_{\sigma{k}}
        \end{equation}
\end{enumerate}
\begin{remark}
    The second point above tries to do the following
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.9\linewidth]{pics/ch6-scanned-notes-1/p9.pdf}
    \end{figure}
\end{remark}

Now we give a precise description of how this Bary center division
will work. We first describe precisely two characteristics of a
simplicial complex.

\begin{defi}[dimension of a simplicial complex]
\nomenclature{dimension of a simplicial complex}{\nomrefpage.}
    The dimension of a simplicial complex $K$ is the maximum of the
    dimensions of all its simplexes.
\end{defi}
\begin{defi}[mesh $\mu(K)$]
\nomenclature{mesh $\mu(K)$}{\nomrefpage.}
    The mesh of a simplicial complex $K$ is the maximum of the
    diameters of all its simplexes.
\end{defi}

As we expected, the mesh (which carries the intuition of the smallness
of one simplicial complex) of a simplicial complex will decrease. The
following lemma confirmed this.

\begin{lemma}
    The sollection of simplexes described above forms a simplicial
    complex. If it is denoted by $K^1$, and is called the first
    barycentric subdivision of $K$, then $K^1$ has the following
    properties:
    \begin{enumerate}
        \item each simplex of $K^1$ is contained in a simplex of $K$;
        \item Their topology are the same: $|K^1|=|K|$
        \item If the dimension of $K$ is $n$, then
            \begin{equation}
                \mu(K^1)\leq \frac{n}{n+1}\mu(K)
            \end{equation}
    \end{enumerate}
\end{lemma}
\begin{proof}
    The proof can be found in page 126 to 127 of \cite{book}. But the
    proof of the last point is not so clear and is illustrated here:
\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{pics/ch6-notes-2/explain1.pdf}
\end{figure}
\end{proof}
\begin{remark}
    Notice that the number $\frac{n}{n+1}<1$, hence after an infinite
    number of Barycentric subdivision, the mesh (averaged diameter) of
    a simplicial complex $K^m$ will goes to zero!
\end{remark}
\subsection{Section 6.3 Simplicial approximation}
\label{sec:Simplicial-approximation}

The purpose here is to approximate any continuous map $f:A\to B$ with
linear map, in a way that is compatible with the triangulations on
both $A$ and $B$.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.4\linewidth]{pics/ch6-notes-2/exp2.png}
\end{figure}
\begin{defi}[Simplicial map]
\nomenclature{Simplicial map}{\nomrefpage.}
    Let $K$ and $L$ be simplicial complexes. A continuous map $s:K\to
    L$ is called simplicial, if it takes simplexes of $K$
    \textit{linearly} to simplexes of $L$. Specifically, being
    \textit{linear} means that:

    For $\forall A\in K$, we have $B=s(A)\in L$. Also, if
    $A=\{v_0,\cdots,v_k\}$, then $B=\{s(v_0),\cdots,s(v_k)\}$.
    Lastly, if $x=\sum_{i=0}^k \lambda_i v_i$ is a a point in $A$, with
    $\lambda_i\geq 0$ and $\sum_{i=0}^k\lambda_i=1$, then
    $$s(x) = \sum_{i=0}^k \lambda_i s(v_i)$$

    In a word, $s$ maps simplexes to simplexes, vertices, to vertices,
    and points linearly to points.
\end{defi}
\begin{remark}
    It is easy to see that the a simplicial map is uniquely determined
    by its value on all the vertices. By this, there can only be a
    finite number of such kind of maps, since there are only a finite
    number of vertices.
\end{remark}

Now we want to find such linear maps to approximate a give function
$f:|K|\to |L|$. Given a point $x\in |K|$, we define the
\nomen{carrier of $f(x)$} to be the unique simplex of $L$ that contain
$f(x)$ as its interior point.

\begin{defi}[Simplicial Approximation]
\nomenclature{Simplicial Approximation}{\nomrefpage.}
    A simplicial map $s:|K|\to |L|$ is called a simplicial approximation of
    $f:|K|\to |L|$ if and only if $s(x)\in$ carrier of $f(x)$,
    $\forall x\in |K|$.
\end{defi}
\begin{ex}
    Trivial examples:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/ch6-notes-2/ex3.png}
\end{figure}
\end{ex}
\begin{ex}
    Now we show that when simplicial approximation does not exits. The
    set up is as follows:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.8\linewidth]{pics/ch6-notes-2/ex4.png}
    \end{figure}
    Hence, there is no simplicial approximation of $f$. But, it will
    be shown below that, if we divide $K$ small enough
    ($K^1,K^2,\cdots$), we will eventually be able to find a
    simplicial approximation of $f$ in some $K^n$.
\end{ex}

\begin{thm}[Simplicial approximation theorem]
    \label{thm:SAT}
    Let $f:|K|\to |L|$ be a continuous map between polyhedras formed
    by simplicial complexes $|K|$ and $|L|$. If $m$ is chosen large
    enough, then there is a simplicial approximation $s:|K^m|\to|L|$
    to $f:|K^m|\to|L|$.
\end{thm}
The proof will use a lemma. Define the \nomen{open star} $v$ of a
simplicial complex $K$, as the union of the interiors of those
simplexes of $K$ having $v$ as a vertex. This union is denoted by
\nomen{$\operatorname{star}(v,K)$}. An example is:
\begin{figure}[H]
    \centering
    \includegraphics[width=1.0\linewidth]{pics/ch6-notes-2/ex5.png}
    \caption{Example of star}
\end{figure}
So it really feels like a star spreading out from $v$. We have

\begin{lemma}
    Vertices $v_0,v_1,\cdots,v_k$ of a simplicial complex $K$ are the
    vertices of a simplex of $K$ if and only if the intersection of
    their open stars is nonempty, i.e. 
    $$\cup_0^k \operatorname{star}(v_i,K)\neq\emptyset$$
\end{lemma}
\begin{proof}
    The proof can be found in page 130 of \cite{book}. It is quite
    trivial.
\end{proof}
Then we have the \textbf{proof of theorem \ref{thm:SAT}}:
\begin{proof}
    The proof is found in page 130 to 131 of \cite{book}. The idea is
    that if for each vertex $u$ of $K$, we can find a vertex $v$ of
    $L$ such that
    \begin{equation}
        f(\operatorname{star}(u,K))\subseteq\operatorname{star}(v,L)
        \label{eq:proof-temp} \tag{*} 
    \end{equation}
    Then the simplicial map $s(u)=v$ will be easily verified to be a
    simplicial approximation of $f$.

    To make such (*) happen, we naturally have to divide $K^m$ small
    enough. We can achieve such small division is guaranteed by the
    Lebesgue lemma \ref{lemma:lebesgue-lemma}, since each subdivision
    makes the averaged diameter smaller.
\end{proof}

\subsection{6.4 The edge group of a complex}
\label{sec:edge-group}

The definition of edge group is so similar to that of fundamental
group (actually, edge group is made to simplify computation of
fundamental group). Therefore, I will not waste too much words here.
One with reasonable intuition could just skim through this section.

The following excerpt from book \cite{book} best captures the idea:

\begin{myquote} \enquote{
We calculated the fundamental groups of one or two spaces in Chapter
5, but our calculations, though efficient for the examples given
there, were rather \textit{ad hoc}. If we agree to work with triangulable
spaces, we can be much more systematic. We shall show how to read off
generators and relationst for the fundamental group from a
triangulation of the space.

Let $X$ be a path-connected triangulable space, take a specific
triangulation $h:|K|\to X$, and replace $X$ by $|K|$ (we are at
liberty to do this since the funda- mental group is a topological
invariant). Now the advantage of a polyhedron $|K|$ is that the
elements of its fundamental group can be represented by loops which
are made up of edges of $K$. Using such 'edge loops' we shall
construct a group, called the edge group of the complex $|K|$, which
can be computed and wh ich is isomorphie to the fundamental group of
$|K|$.
} \end{myquote}

The edge pathes $v_0,v_1,\cdots,v_k$, edge loops
$v,v_1,\cdots,v_{k-1},v$ are defined similar to pathes and loops.
The equivalence relation between edge loops are defined as represented
by the following drawing:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/ch6-notes-2/edge-group1.png}
\end{figure}

The equivalence class of edge pathes and edge loops are denoted
$\{v_0,v_1,\cdots,v_k\}$, and $\{v,v_1,\cdots,v_{k-1},v\}$,
respectively. The multiplication of edge loops is defined simply as
\begin{equation}
    \{v,v_1,\cdots,v\}\{v,w_1,\cdots,v\} =
    \{v,v_1,\cdots,v,w_1,\cdots,v\}
\end{equation}

It is trivially verified that such edge loops form a group, called the
\nomen{edge group based at $v$, $E(K,v)$}. It is also trivially
verified that such group does not depend on the choice of base point
$v$, since $|K|$ is path-connected.

\begin{thm}
    $E(K,v)$ is isomorphie to $\pi_1(|K|,v)$
\end{thm}
\begin{proof}
    The proof can be found on page 133 of \cite{book}. The essential
    idea is that an edge loop is just a loop. And a loop is just a
    continuous map $f:I\to |K|$, we can easily use the simplicial
    approximation $s:I^m\to |K|$ of $f$ (with appropriately $m$), to
    find a corresponding edge loop.

    The uniqueness of the loop corresponding to an edge loop is found
    again by treating $L:=I^2$ as a simplicial complex and apprximate
    again the homotopy $F:L\to |K|$. The resulted $S$ will again gives
    us a equivalence relation to the constant edge loop. (Note, we
    have to use the fact the simplicial maps preserves the equivalence
    relation between edge pathes). The following pictorially
    produces the proof:
    \begin{figure}[H]
        \centering
        \includegraphics[width=1\linewidth]{pics/ch6-notes-2/L.png}
    \end{figure}
\end{proof}

>>>>

\subsection{Section 6.2 Barycentric Division}
\label{sec:Barycentric-Division}
Now we begin to make the triangulation more and more detailed. We
start from a simplicial complex $K$ and constructing a new simplicial
complex $K_1$, such that
\begin{enumerate}
    \item $|K|=|K_1|$, homeomorphically.
    \item diameter $|K_1|<$ diameter $|K|$.
\end{enumerate}
The general construction method is called Barycentric division,
visualized as:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/ch6-scanned-notes-1/8.pdf}
\end{figure}
We first define a concept
\begin{defi}[$\hat{A}$ Bary center]
\nomenclature{$\hat{A}$ Bary center}{\nomrefpage.}
Let $A$ be a simplex, then the bery center of $A$, denoted $\hat{A}$,
is the point:
\begin{equation}
    \hat{A}:= \frac{1}{k+1}(v_0+\cdots+v_k)
\end{equation}
\end{defi}
Then, the general process is:
The new space $K_1$ is such that
\begin{enumerate}
    \item The vertices of $K_1$ are bary centers of \textbf{ALL}
        simplexes of $K$. Note the the bary center of a $0$-simplex
        is just itself. So in general, $K_1$ has more vortices of
        $K$.
    \item A collection $\hat{A}_0,\cdots,\hat{A}_k$ of such bary
        centers will form a vortex of a $k$-simplex in $K_1$ if and
        only if there is a permutation $\sigma$ of $\{0,1,\cdots,k\}$
        such that
        \begin{equation}
            A_{\sigma{0}}< A_{\sigma{1}} < \cdots < A_{\sigma{k}}
        \end{equation}
\end{enumerate}
\begin{remark}
    The second point above tries to do the following
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.9\linewidth]{pics/ch6-scanned-notes-1/p9.pdf}
    \end{figure}
\end{remark}

Now we give a precise de
